---
title: "Project 1"
date: "10/1/2021"
output: html_document
---

## Peer-graded Assignment: Activity monitoring data
### 0. Packages needed:

```{r, message=FALSE}
library(ggplot2)
library(tidyverse)
```


### 1. Reading the dataset:
```{r eval=FALSE}
download.file("https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip", destfile = "./Reproducible research/data.zip")
                
unzip("./Reproducible research/data.zip")
```

```{r}
data <- read.csv("./Reproducible research/activity.csv")
```


#### Processing the data:

```{r }
data$date<-as.POSIXct(data$date, format="%Y-%m-%d")
```


### 2. Histogram of the total number of steps taken each day

```{r error=TRUE}
dailySteps <- aggregate(steps ~ date, data = data, sum, na.rm = TRUE)
```

```{r}
hist(dailySteps$steps,xlab = "Day", 
     main="Total steps taken each day", col = "paleturquoise", breaks = 40)
```

### 3. Mean and median number of steps taken each day

```{r}
mean(dailySteps$steps)
median(dailySteps$steps)
```


### 4. Time series plot of the average number of steps taken

```{r}
avgSteps <- aggregate(steps ~ interval, data, mean)
ggplot(avgSteps, aes(x = interval,y = steps)) + geom_line() + 
        xlab("5-Minute interval") + ylab("Average steps") + theme_classic()
```

The 5-minute interval, that contains the maximum number of steps on average across all the days in the dataset,  is:

```{r}
avgSteps$interval[avgSteps$steps == max(avgSteps$steps)]
```

### 5. The 5-minute interval that, on average, contains the maximum number of steps

```{r}
avgSteps[which.max(avgSteps$steps),]
```

The 835th interval contains the average maximum number of steps: 206.16

### 6. Code to describe and show a strategy for imputing missing data

```{r}

#Number of missing values

sum(is.na(data$steps))
```

__Strategy to fill the missing values:__ 

```{r}
# Replacing missing data with the mean, and creating a new data set

newData <- data %>%
          mutate(steps
          = replace(steps, is.na(steps),
                   mean(steps, na.rm = TRUE)))
```


### 7. Histogram of the total number of steps taken each day after missing values are imputed

```{r}
newDailySteps <- aggregate(steps ~ date, data = newData, sum, na.rm = TRUE)
hist(newDailySteps$steps,xlab = "Day", 
     main="Total steps taken each day", col = "paleturquoise", breaks = 40)
```

__New mean and median of the data__:

```{r}
mean(newDailySteps$steps)
median(newDailySteps$steps)
```

There is a very small difference between the original mean and median and the new one.

### 8. Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends


``` {r, echo = TRUE}
newData$days <- weekdays(newData$date)
newData$weekday <- ifelse(weekdays(newData$date) %in% c("sábado", "domingo"), "weekend", "weekday")
newAvgSteps <- newData %>% group_by(weekday, interval) %>% summarise(meanSteps = mean(steps))

ggplot(newAvgSteps, aes(x = interval, y = meanSteps)) +
  geom_line() + facet_grid(weekday ~.) + 
  xlab('5-Minute interval') + ylab('Average steps') +
  theme_classic()

```

We can see some differences in activity patterns, during the weekends the activity seems constant during the day, on the weekdays the activity seems lower after a spike early in the day.
